---
title: "Simulation for poisson ash using normal mix prior"
author: "Yusha Liu"
header-includes:
   - \usepackage{amsmath}
   - \usepackage{bm}
   - \usepackage{bbm}
   - \usepacakge{mathtools}
date: "2020-06-09"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction
Run a simple simulation using poisson ash, which places a normal mixture prior on $\lambda_i$. The data are generated following http://dattahub.github.io/GHstancodes.

### Model
\begin{align}
y_i & \sim Pois(s_i \lambda_i) \\
\beta_i & = \log(\lambda_i) \sim \sum_k \pi_k N(\mu, \sigma_k^2) \\
\end{align}
Letting $\tilde{\beta}_i = \beta_i - \mu$, we approximate the posterior $p(\tilde{\beta}_i \mid y_i, \mu, \pi)$ with a mixture of normal $\sum_k \zeta_{ik} N(m_{ik}, v_{ik}^2)$ using variational Bayes. 

```{r warning=FALSE, message=FALSE}
library(ggplot2)
theme_set(theme_bw())
library(plyr)
library(dplyr)
library(reshape2)
source("code/util.R")
```

### Simulate Poisson data
```{r warning=F}
set.seed(786)
n = 200; w = 0.9
idx = rep(1,n)
lambdasparse = rep(0,n)
for (i in 1:n)
{
  if(i<=round(n*w)){
    lambdasparse[i]<- 0.1
    idx[i] <- 0}
  else {lambdasparse[i] <- 10}}
y <- rpois(n,lambdasparse); s <- rep(1, n)
```

### Fit Poisson ash
```{r warning=F}
sigma2 <- c(1e-6, 1e-5, 1e-4, 4e-4, 1e-3, 4e-3, 1e-2, 2e-2, 4e-2, 8e-2, 0.16, 0.32, 0.64, 1, 2, 4, 8, 16)

RUNTIME = c()
start_time = proc.time()
fit <- pois_ash(y, s, sigma2, init=NULL, maxiter=200, tol=1e-6, verbose=FALSE)
lambda.pm <- rowSums(fit$Ez * exp(fit$mu + fit$m + fit$v2/2))
runtime = proc.time() - start_time
```

### ELBO and runtime
```{r warning=F}
## ELBO plot over iteration
plot(fit$ELBO, xlab = "niter", ylab = "elbo")

## running time
runtime
```

### Check results
```{r}
## estimated pi
fit$pi

## compare posterior mean with the observed count and true lambda
post.data = rbind(data.frame(type = "true lambda", values=lambdasparse, x=seq(1:n)),
                  data.frame(type = "observation", values=y, x=seq(1:n)),
                  data.frame(type = "posterior mean", values=lambda.pm, x=seq(1:n)))

ggplot(post.data, aes(x=x, y=values, group=type, colour=type)) + 
  geom_point(aes(colour = type),size = 1) + ylab("Possion ash") + xlab(expression(y))+facet_grid(type~.,scales="free_y") + ylim(range(y))
```
